### 2013-02-05 doc/xps2013/proposal.tex
some notes on the research section

\subsection{Improved Networking layer}

@@

Hierarchy will reduce scaling requirements from \BigO{n^2} threads to \BigO{n log(n)}.  Of course, it takes what used to be a single network hop for communication and turns it into \BigO{log(n)} hops.  More threads will be required to overcome this increased latency, but we believe only \BigO{log(n)} more, hence a real scaling benefit is achievable with the a virtual hierarchical network topology.

@@

This interface suffers from multiple shortcomings.  The biggest of which is that ultimately, moving data to and from a remote node constructively requires higher level semantic operations than simple reads and writes.  These higher level semantics can, of course, be synthesized from the underlying read and write primitives, but it wastes round trips and synchronization time.  One possible alternative solution is to borrow ideas from Active Messages~\cite{active-messages} and enable remote nodes to dispatch, essentially, user-level function calls.  The challenge here is to make those calls interoperate securely with a modern operating system and multicore system and be efficient.  \comment{This seems weak.  Need some more ideas here.  -Mark}

The second challenge is the virtual to physical address translation on the cards requires the operating system to pin translations.  This removes paging flexibility from the operating system and allows user processes to (potentially unfairly) grab ahold of physical memory.  In the long run, the solution here likely lays at the intersection of what progress is occurring in virtual memory translation for I/O with virtual machines and RDMA.  Our goal will be to carefully study the I/O virtualization proposals put forth by major processor manufactures and devise solutions that enable RDMA to capitalize on these developments.  \comment{weak?  -M}

@@

>>>>

\comment{I think we should pitch what have been done so far as essays and a bit part of the research effort is putting everything together. -Luis}

\subsection{Networking}

\comment{This research item	is important but right now it sounds really really
specific. You almost need to know the current code to understand it. -Luis}

\comment{Ins't the argument here something like: If we did aggregation on a
per-destination basis only, as the number of nodes scales, it takes more
concurrency (or more time) to aggregate messages. We will address this by
doing aggrregation hierarchically, choosing representative end-points along
the way to increase agggregation opporunitites at the cost of more
communucation hops (making an analogy with changing trains here). -Luis}

@@

% this paragraph is great!!
Hierarchy will reduce scaling requirements from \BigO{n^2} threads to \BigO{n log(n)}.  Of course, it takes what used to be a single network hop for communication and turns it into \BigO{log(n)} hops.  More threads will be required to overcome this increased latency, but we believe only \BigO{log(n)} more, hence a real scaling benefit is achievable with the a virtual hierarchical network topology.

@@

This interface suffers from multiple shortcomings.  The biggest of which is that ultimately, moving data to and from a remote node constructively requires higher level semantic operations than simple reads and writes.  These higher level semantics can, of course, be synthesized from the underlying read and write primitives, but it wastes round trips and synchronization time.  One possible alternative solution is to borrow ideas from Active Messages~\cite{active-messages} and enable remote nodes to dispatch, essentially, user-level function calls.  The challenge here is to make those calls interoperate securely with a modern operating system and multicore system and be efficient.  \comment{This seems weak.  Need some more ideas here.  -Mark. Luis: Isn't the problem that the app runtime should be able to determine how the RDMA ops should be done? Maybe add examples here like accessing a hash-table, doing collective operations on remote data etc (min, max, avg)? How about delegating simple operations to the network card? Maybe argue that getting the attention of the remote processor for these operations is too costly?}

The second challenge is the virtual to physical address translation on the cards requires the operating system to pin translations.  This removes paging flexibility from the operating system and allows user processes to (potentially unfairly) grab ahold of physical memory.  In the long run, the solution here likely lays at the intersection of what progress is occurring in virtual memory translation for I/O with virtual machines and RDMA.  Our goal will be to carefully study the I/O virtualization proposals put forth by major processor manufactures and devise solutions that enable RDMA to capitalize on these developments.  \comment{weak?  -M If you are 100\% sure this is true, then it is not weak, but if we are wrong, then people can shoot down the proposal saying we know nothing about how RDMA is done today...}

@@

\comment{mostly a copy and paste of the pomace?}
